---
output:
  md_document:
    variant: markdown_github
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# ncapi

The goal of ncapi is to provide a foundation interface to the NetCDF API that transcends the three overlapping and patchily supported independent packages 
`ncdf4`, `RNetCDF` and `rhdf5`. 

The need for this is discussed here:  

http://rpubs.com/cyclemumner/293536

The crux is cross-platform support for NetCDF class, NetCDF-4 and HDF5-compatibles, Thredds servers, and sources
with groups and compound types. 

The first useable version of this package would replace the use of `RNetCDF` in https://github.com/hypertidy/ncmeta. A key need is to have
consistent cross-platform support for NetCDF-4 and -3, Thredds servers, compression options and MPI options. 

Later versions might be used by `RNetCDF` or `ncdf4`, 

# Interested? Great!!

I have no idea what I'm doing so any assistance, ideas, testing, encouragement is welcome.

In particular I need help with: 

* basic C++, dealing with numbers and characters passed with the NetCDF API
* higher level design and C++ implementations to reduce the maintenance footprint
* specific workflows and sources to test

Source of interest include: 

* classic NetCDF 3 files
* NetCDF 4 files (the ones that can have *groups* and/or *compound-types*)
* HDF5 files
* Thredds, DODS, OpenDAP servers
* files with pathologies, e.g. character string coordinate values, unused dimensions, broken entities
* complex multi-dimensional models
* NetCDF sources that use compound types for examples different to L3BIN NASA ocean colour or KEA
* anything I haven't thought of that you know about ...


# set up notes

* General system requirements: 

https://github.com/mdsumner/nectar/blob/master/r-spatial.sh

* package specific details See DESCRIPTION, ncapi-package.r, src/init.c, src/Makevars

* Register routines, new in R 2017:  https://ironholds.org/registering-routines/

http://dirk.eddelbuettel.com/blog/2017/04/30/#006_easiest_package_registration

* update registration with `tools::package_native_routine_registration_skeleton("../ncapi", "src/init.c",character_only = FALSE)`

* Configure Rcpp 1) must import something  '@importFrom Rcpp evalCpp' 2) useDynLib in roxygen


# raw examples

```{r L3bin}
library(ncapi)
f_l3b <- system.file("extdata", "oceandata", "S2008001.L3b_DAY_CHL.nc", package = "ncapi")
con <- Rnc_open(f_l3b)
groupids <- Rnc_inq_grps(con)
Rnc_inq_grpname(groupids[1])
lapply(Rnc_inq_grps(con), Rnc_inq_grpname)
Rnc_close(con)
```

Get a large-ish summary of what is in the file (very WIP). 

```{r}
example(Rnc_inq)
```

Simple get group names. 

```{r func}
get_groups <- function(x, check_exists = TRUE) {
  if (check_exists) stopifnot(file.exists(x))
  con <- Rnc_open(x)
  on.exit(Rnc_close(con), add = TRUE)
  groupids <- Rnc_inq_grps(con)
  cat(sprintf("returning %i group names\n", length(groupids)))
  names <- unlist(lapply(groupids, Rnc_inq_grpname))
  
  if (length(names) < 1) {
    names <- "<no groups found>"
    groupids <- NA_integer_
  }
   tibble::tibble(group_id = groupids, group_name = names, source = basename(x), access = Sys.time())
}
files <- list.files(system.file("extdata", package = "ncapi"), recursive = TRUE, pattern = "nc$", full.names = TRUE)
d <- dplyr::bind_rows(lapply(files, get_groups), .id = "source")
print(d)
```

# Thredds

(These sources don't have groups so we need some more functionality to see anything happen).


```{r}
u1 <- "http://tds.hycom.org/thredds/dodsC/GLBa0.08/latest/2d"
get_groups(u1, check_exists = FALSE)


u2 <- "https://oceandata.sci.gsfc.nasa.gov:443/opendap/MODISA/L3SMI/2016/001/A20160012016032.L3m_R32_SST_sst_9km.nc"
get_groups(u2, check_exists = FALSE)
```


Investigate them a little more deeply. 

```{r}
con <- Rnc_open(u1)
Rnc_inq_grps(con)

## notice how the connection is the group ID in the degenerate case
summ1 <- Rnc_inq(con)
Rnc_close(con)


con <- Rnc_open(u2)
summ2 <- Rnc_inq(con)
Rnc_close(con)
```


Content summary of hycom **GLBa0.08**. 

```{r}
print(summ1)
```

Content summary of **A20160012016032.L3m_R32_SST_sst_9km**.

```{r}
print(summ2)
```


# Archaeology

This project is a progression from past attempts to make sense of this space. 

* https://github.com/mdsumner/anc
* https://github.com/hypertidy/rancid
* https://github.com/hypertidy/ncdump
* http://lists.r-forge.r-project.org/pipermail/rcpp-devel/2016-December/009485.html
* https://github.com/RConsortium/wishlist/issues/3

# Code of conduct

Please note that this project is released with a [Contributor Code of Conduct](CONDUCT.md). By participating in this project you agree to abide by its terms.

